Economic Evaluation of Mis-speficied Geographically Weighted Regression Models for Site-specific Nitrogen Management

### 2023-06-23 Writing/manuscript_gwr.rmd

However, deploying sophisticated sensors to gather field characteristics data is also costly, and the effects of their generation and use on farming profitability remains in question [@gardner2021economic]. Recognizing the significant costs of obtaining useful, high-resolution soil and field properties data, @trevisan2021spatial demonstrated a method of providing site-specific management recommendations without such data. They estimated site-specific yield response functions by applying GWR techniques on data from OFPEs [@bullock2019data], and reported dramatic results, estimating an average potential gain of \$65 ha$^{-1}$ from applying seed and N at optimal uniform rates (instead of at the rates the farmers would have chosen if not participating in the research), and an additional potential gain of \$58 ha$^{-1}$ when changing from optimal uniform application rates to optimal site-specific application rates. This is rather surprising given the often-reported poor statistical accuracy of GWR estimates [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation] and also much lower value of variable rate application over uniform rate application reported in previous studies [@bullock1998does; @thrikawala1999economic; @koch2004economic]. @rakshit2020 also applied GWR to on-farm experiment data. However, while they did estimate site-specific yield response functions, they did not derive site-specific economically optimal nitrogen rates based on them, and thus did not conduct economic analysis, unlike @trevisan2021spatial. @evans2020assessment explored using GWR with a linear model specification to detect low- and high-response areas, but like @rakshit2020, they also did not estimate site-specific economically optimal nitrogen rates.

@@

The second source of over-estimation bias is endemic to any study that evaluates the value of variable rate application (VRA), not just @trevisan2021spatial. In general, researchers take the following steps to estimate the value of VRA using OFPE data [e.g., @anselin2004spatial; @swinton1998evaluating; @bullock2002adding; @liu2006site]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIRs) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the step 3's assumption that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provided an example at step (3) by assuming that their GWR estimates provided "true" site-specific yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true site-specific yield response functions. But, inaccurate estimation of yield response function in this methodology can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions, such as GWR.

The primary objective of this study is examine the degree and direction of bias in the estimation of the economic value of GWR-based site-specific nitrogen rate recommendations (or "variable rate application", VRA) when GWR model is specified as quadratic when the true yield response function follows the quadratic-plateau functional form. We evaluate two sets of GWR quadratic specifications used in @rakshit2020 and @trevisan2021spatial. Both of them assumed a quadratic response of yield to nitrogen rate ($N$). However, while @trevisan2021spatial estimated the coefficient on $N^2$ as a global parameter, @rakshit2020 estimated it locally for each point. At first, it appears that the specification used by @rakshit2020 is superior since it is more flexible than the specification used by @trevisan2021spatial. However, the flexibility can be a disadvantage rather than an advantage because the estimation of economically optimal nitrogen rate (EONR) is highly sensitive to the coefficient estimates on $N^2$, and GWR often provides inaccurate local estimation of that coefficient. It is critical to keep in mind that it is not our objective to measure the bias in the coefficient estimation by GWR, unlike [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation]. Rather, we are interested in the economic implications of the statistical inaccuracy in estimating the coefficients of the model in this study. As part of identifying bias in the estimation of the economic value of VRA based on the GWR approaches, we evaluate their actual values. Finally, we also evaluate their relative performances compared to uniform rate application (URA) based on shape-constrained additive model approach. It should be noted that the economical optimization in this study only involves farmers' private profit, while the environmental externalities (such as fertilizer runoff and leaching) are not considered.

We pursue the objective by conducting Monte Carlo simulations in which we first generate and then analyze stylized OFPE data. We cannot test which model specification results in more profitable site-specific nitrogen rate recommendations using real data because it is impossible to observe the true yield response functions using real-world data. For the same reason, we cannot test whether @trevisan2021spatial over-estimates the actual value of VRA, either. But in the Monte Carlo simulations we specify yield response functions and generate stylized data from those functions, which enables us to quantify in the context of the simulations the value of GWR and the bias in GWR-based estimation of the value of VRA. Monte Carlo simulations are commonly used in this way to gain insights into the statistical and economic properties of trial designs and estimations methods [@mckinion2001analysis; @gebbers2010application; @bullock2020value; @alesso2021design].

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. Our simulation assume that each cell $(j, i)$'s "true" yield response function is assumed to follow the commonly-used quadratic-plateau functional form:

@@

It is worthy of mentioning that, in most previous studies of model comparisons the model performances were predominantly measured by statistical accuracy of parameter estimates (such as bias, mean squared error, type I error, confidence intervals, and so on), for either the GWR models [@lu2017geographically, @harris2019simulation] or the econometric models in general [@christopeit2006local, @su2016identifying]. While the statistical measures are fundamental criteria to evaluate model performances, what matters more in economic practices are whether and how much those statistical accuracy can translate into the ultimate economic values. Particularly, the magnitudes of this translation may depend on the application scenarios. This study differs from previous literature in that it reports the economic value of specific econometric models (GWR) with model functional form misspecification (quadratic estimation while the truth is quadratic-plateau) when being used in specific applications (corn site-specific nitrogen management) and under specific conditions (normal-sized production field, various market price scenarios, etc.) through a simulation-based approach.

>>>>

However, deploying sophisticated sensors to gather field characteristics data is also costly, and the effects of their generation and use on farming profitability remains in question [@gardner2021economic]. Recognizing the significant costs of obtaining useful, high-resolution soil and field properties data, @trevisan2021spatial demonstrated a method of providing site-specific management recommendations without such data. They estimated site-specific yield response functions by applying GWR techniques on data from OFPEs [@bullock2019data], and reported dramatic results, estimating an average potential gain of \$65 ha$^{-1}$ from applying seed and N at optimal uniform rates (instead of at the rates the farmers would have chosen if not participating in the research), and an additional potential gain of \$58 ha$^{-1}$ when changing from optimal uniform application rates to optimal site-specific application rates. This is rather surprising given the often-reported poor statistical accuracy of GWR estimates [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation] and also much lower value of variable rate application (VRA) over uniform rate application (URA) reported in previous studies (cite). @rakshit2020 also applied GWR to on-farm experiment data. However, while they did estimate site-specific yield response functions, they did not derive site-specific economically optimal nitrogen rates based on them, and thus did not conduct economic analysis, unlike @trevisan2021spatial. @evans2020assessment explored using GWR with a linear model specification to detect low- and high-response areas, but like @rakshit2020, they also did not estimate site-specific economically optimal nitrogen rates.

@@

The second source of over-estimation bias is endemic to any study that evaluates the value of variable rate application, not just @trevisan2021spatial. In general, researchers take the following steps to estimate the value of VRA using OFPE data [e.g., @anselin2004spatial; @swinton1998evaluating; @bullock2002adding; @liu2006site]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIRs) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the step 3's assumption that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provided an example at step (3) by assuming that their GWR estimates provided "true" site-specific yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true site-specific yield response functions. But, inaccurate estimation of yield response function in this methodology can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions, such as GWR [@czarnota2015evaluating;].

The primary objective of this study is examine the degree and direction of bias in the estimation of the economic value of GWR-based site-specific nitrogen rate recommendations (or "variable rate application", VRA) when GWR model is specified as quadratic when the true yield response function follows the quadratic-plateau functional form. We evaluate two sets of GWR quadratic specifications used in @rakshit2020 and @trevisan2021spatial. Both of them assumed a quadratic response of yield to nitrogen rate ($N$). However, while @trevisan2021spatial estimated the coefficient on $N^2$ as a global parameter, @rakshit2020 estimated it locally for each point. At first, it appears that the specification used by @rakshit2020 is superior since it is more flexible than the specification used by @trevisan2021spatial. However, the flexibility can be a disadvantage rather than an advantage because the estimation of economically optimal nitrogen rate (EONR) can be highly sensitive to the coefficient estimates on $N^2$, and GWR often provides inaccurate local estimation of that coefficient. It is critical to keep in mind that it is not our objective to measure the bias in the coefficient estimation by GWR, unlike [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation]. Rather, we are interested in the economic implications of the statistical inaccuracy in estimating the coefficients of the model in this study. As part of identifying bias in the estimation of the economic value of VRA based on the GWR approaches, we evaluate their actual values. Finally, we also evaluate their relative performances compared to uniform rate application (URA) based on shape-constrained additive model approach. It should be noted that the economical optimization in this study only includes farmers' private profit, while the environmental externalities (such as fertilizer runoff and leaching) are not considered.

We pursue the objective by conducting Monte Carlo simulations of 1000 rounds in which we first generate and then analyze stylized OFPE data. We cannot test which model specification results in more profitable site-specific nitrogen rate recommendations using real data because it is impossible to observe the true yield response functions using real-world data. For the same reason, we cannot test whether @trevisan2021spatial over-estimates the actual value of VRA, either. But in the Monte Carlo simulations we specify yield response functions and generate stylized data from those functions, which enables us to quantify in the context of the simulations the value of GWR and the bias in GWR-based estimation of the value of VRA. Monte Carlo simulations are commonly used in this way to gain insights into the statistical and economic properties of trial designs and estimations methods [@mckinion2001analysis; @gebbers2010application; @bullock2020value; @alesso2021design].

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. Our simulation assumes that each cell $(j, k)$'s "true" yield response function is assumed to follow the commonly-used quadratic-plateau functional form:

@@
